---
title: "Exploring Discretionary Accruals"
author: "Joachim Gassen"
institute: "TRR 266 Accounting for Transparency"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: "9pt"
output: 
  beamer_presentation: 
    latex_engine: xelatex # pdflatex creates rastered fonts
    slide_level: 3

        
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \input{beamer_theme_trr266.sty}
---


```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(
  cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE
)
library(tidyverse)
library(kableExtra)
load("../output/results.rda")
```


# Motivation

### Discretionary Accruals

- Are used in tons of papers
- As measurement constructs for almost everything
  - Earnings Management
  - Earnings Quality
  - Audit Quality
  - Management Quality
- But how do they look like?


### Data

```{r tab_descriptives_sample, results="asis"}
desc_rnames <- paste0("\\textit{", var_names$label, "}")

t <- tab_desc_stat
rownames(t$df) <- desc_rnames
names(t$df)[c(5,7)] <- c("25 \\%", "75 \\%")
kable(
  t$df, digits = c(0, 3, 3, 3, 3, 3, 3, 3), format = "latex",
  format.args = list(decimal.mark = ".", big.mark = ",", scientific=FALSE),
  booktabs = TRUE, escape = FALSE, linesep = ""
) -> kab_latex
lat_tab <- unlist(strsplit(kab_latex, "\n"))
latex_tab <- c(
  "\\begin{table}[!htbp]",  
  "\\resizebox{\\textwidth}{!}{",
  "\\begin{threeparttable}",
  lat_tab[2:(length(lat_tab) - 1)],
  "\\end{tabular}",
  "\\begin{tablenotes}[flushleft]",
  "\\setlength{\\labelsep}{0pt}",
  sprintf(
    paste(
      "\\item Note: The data is obtained from Compustat U.S. as provided",
      "by WRDS. The sample covers the period %d to %d and %s unique firms."
    ),
    min(smp_da$fyear), max(smp_da$fyear), 
    format(length(unique(smp_da$gvkey)), big.mark = ",")
  ),
  "\\end{tablenotes}",
  "\\end{threeparttable}",
  "}",
  "\\end{table}"
)

cat(paste(latex_tab, collapse = "\n"))  
```


### Distribution of Discretionary Accruals over time - unrestricted sample

```{r BPlotFULL}
fig_boxplot_full
```


### Distribution of Discretionary Accruals over time - analysis sample

```{r BPlotSMP}
fig_boxplot_smp
```


### Correlation of Modified Jones with Dechow and Dichev DA

```{r CorrVizz}
fig_scatter_md_dd
```


### Association of Dechow and Dichev DA with firm size

```{r AssDDSize}
fig_scatter_dd_lnta
```


### No presentation without correlation table...

``` {r CorrTable, results="asis"}
tab <- tab_corr
lat_tab <- unlist(strsplit(tab$kable_ret, "\n"))
lat_tab[7:13] <- str_replace(
  lat_tab[7:13], fixed(var_names$var_name), var_names$label
)
lat_tab[4] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
lat_tab[6] <- "\\hline\\\\[-1.8ex]"
lat_tab[14] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
latex_tab <- c(
  "\\begin{center}",
  "\\resizebox*{\\textwidth}{!}{",
  "\\begin{tabular}{lccccccc}",
  lat_tab[4:14],
  "\\end{tabular}}",
  "\\end{center}"
)
cat(paste(latex_tab, collapse = "\n"))
```


### ... and regression table

``` {r RegTable, results="asis"}
tab <- tab_regression$table
tab[12] <- paste(
  "\\\\[-1.8ex] &", var_names$label[1], 
  "&", var_names$label[2], "\\\\"
)
esc_var_names <- str_replace_all(var_names$var_name, fixed ("_"), "\\_")
reg_names <- tibble(
  var_name = esc_var_names[c(3, 5:7)],
  label = var_names$label[c(3, 5:7)]
)
for (i in seq(15, 24, by = 3)) {
  pos <- (i-12)/3
  tab[i] <- str_replace(
    tab[i], fixed(reg_names$var_name[pos]), reg_names$label[pos]
  )
}
latex_tab <- c(
  "\\begin{center}",
  "\\resizebox*{!}{0.8\\textheight}{",
  "\\begin{tabular}{lcc}",
  tab[8:36],
  "\\end{tabular}}",
  "\\end{center}"
)
cat(paste(latex_tab, collapse = "\n"))
```


### Conclusion

- Discretionary accruals are simply wonderful! 
- With these distributional properties and the power of large samples they will almost always be significant in one direction or the other regardless the setting or controls
- Use them for every economic construct under the sun!







